Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2013-11-27 paper/paper.tex
a few pill edits Tested everything on collins.

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-300-50-0-0-1500.pdf}
  \caption{Images of the concentration of each species in a natural
    pill-shaped bacterium as a function of time, with darker regions
    indicating higher concentration.  Frames from an animation showing
    a series of contour plots of the concentration of proteins in
    different stages over time.  The order of frames is such that
    individual MinD proteins begin at the bottom of the plot (in the
    MinD:ATP state in the cytoplasm), and progress upward until they
    reach the MinE:MinD:ATP membrane-bound complex.  At that point,
    they will spontaneously dissociate into cytoplasmic MinE (the top
    row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.}
  \label{image-p}
\end{figure*}

@@

In the beginning, the cytoplasmic MinD:ATP is roughly evenly
distributed throughout the cell with close to its peak density.  There
is a high concentration of MinD on the membrane (in both its
membrane-bound states) at the lower end of the cell, and there is also
cytoplasmic MinD:ADP concentrated in the lower end, which has recently
left the cell membrane.   There is relatively little cytoplasimic
MinE, as it is primarily bound to the lower end of the membrane.

After ??? seconds, very little MinD remains bound to the membrane, and
there is a high concentration of cytoplasmic MinE in the lower half of
the cell.  In the mean time MinD has begun binding in significant
quantities to the upper half of the cell membrane, resulting in a
diminishing of the concentration of cytoplasmic MinD:ATP in the upper
half of the cell.  At this stage, we begin to see the formation of the
``MinE ring'' in the center of the cell---visible here in the plot of
the concentration of membrane-bound MinE:MinD:ATP as two lines
representing a cross section of the ring.  This ring forms as the
cytoplasmic MinE reacts with the membrane-bound MinD:ATP, and leads to
the removal of MinD from the membrane.

From ??? seconds to ??? seconds, we see the MinE ring progress
upwards, as the cytoplasmic MinE diffuses in that direction,
progressively removing the membrane-bound MinD from eventually all but
the very top of the cell membrane.  During this process, the MinD:ATP
bound to the top of the cell membrane is attracting increasing
quantities of MinD:ATP from the cytoplasm.  At ??? seconds, the cell
is in the mirror image state from which it started, and the period is
half-complete.  The second half of the period is a mirror image of the
process just observed, as MinD is driven from the membrane at the top
of the cell, begins binding to the lower half of the cell and a MinE
ring pushes it back down to lower end of the cell.

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-300-50-0-0-1500}
  %\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above) and a
    3$\micron$ by 1$\micron$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

\davidsays{We should make Fig. \ref{box-p} and Fig. \ref{image-p} plot
  for precisely the same times, that will make it way easier to read
  the two plots.}  While Fig.~\ref{image-p} is helpful for visualizing
the process occuring in the Min system, it is does not show the
relative number of proteins in each state.  To visualize this
information, we plot in Fig.~\ref{box-p} the fraction of MinD found in
each stage of the cycle in each of three sections of the cell---top,
middle and bottom---as a function of time over one period.  The time
axis is the same as in Fig.~\ref{image-p}.  Looking at the evolution
of the latter, we see an initial high concentration of cytoplasmic
MinD:ATP, as the this protein diffuses into the polar region, followed
by a dip in this and large spike in membrane bound MinD:ATP, as the
protein clusters onto the polar cell wall, followed by a spike in the
memebrane bound MinE:MinD:ATP stage of protein, together with a slight
rise in cytoplasmic MinD:ADP as the MinD dettaches from the wall.

@@

\fixme{our 5 length cells are labeled as 4 cells in the file labels,
  since 4 is the length of the cylinder. In the text I'll refer to the
  5 length cells (file says 4) as 5 length cells.  What huang's paper
  calls 4 length cells will be labeled by our files as 3 length cells,
  but will be refered in our text as 4 length cells}.  We can see a
clear difference in period between the $5\mu$, $4\mu$, and $2\mu$
length cells.  The oscillation periods for our $4\mu$ show excellent
agreement with the simulations of Huang et all
\cite{huang2003dynamic}.  This is to be expected since we begin our
simulations with their reported wild type concentrations of MinD and
MinE proteins. The periods for the different pill lengths are shown in
the table below.  We can see a direct correltation between size and
period, as the proteins must diffuse further in a longer cell before
they accumulate on the opposite wall.

>>>>

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-300-50-0-0-1500.pdf}
  \caption{Images of the concentration of each species in a natural
    pill-shaped bacterium as a function of time, with darker regions
    indicating higher concentration.  The cells pictured are
    $4\micron$ in length, measured from end to end.  Frames from an animation showing
    a series of contour plots of the concentration of proteins in
    different stages over time.  The order of frames is such that
    individual MinD proteins begin at the bottom of the plot (in the
    MinD:ATP state in the cytoplasm), and progress upward until they
    reach the MinE:MinD:ATP membrane-bound complex.  At that point,
    they will spontaneously dissociate into cytoplasmic MinE (the top
    row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.}
  \label{image-p}
\end{figure*}

@@

At $t=0$ in Figs.~\ref{image-p}~and~\ref{box-p}, the cytoplasmic MinE
is concentrated in the lower portion of the cell, and is diffusing
upward.  The MinD protein is initially concentrated on the membrane in
the upper portion of the cell.  At the boundary between the
cytoplasmic MinE and the membrane-bound MinD, there is a ring of the
membrane-bound MinE:MinD complex, which is moving towards the upper
end of the cell, removing the MinD from the membrane as it goes.  The
MinD is removed from the membrane in its MinD:ADP form, which begins
to accumulate at the upper end of the cell, peaking peaking at around
10 seconds.

At 15 seconds, the membrane-bound MinD:ATP has been essentially
removed from the top end of the cell, and MinD:ATP has begun to bind
to the membrane in the the lower half of the cell.  At this stage,
there is a high concentration of cytoplasmic MinE at the top of the
cell, and by 16 seconds we begin to see the formation of a MinE ring
just below the center of the cell.  At 18 seconds, the cell has
reached its original state, with MinD:ATP bound to the membrane on the
lower third of the cell, a high concentration of cytoplasmic MinE in
the upper half of the cell, and a MinE ring pushing on the
membrane-bound MinD:ATP.

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-300-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a natural pill shaped cell.  The cell is the same as is
    shown in Fig.~\ref{image-p}.  The time axis is also identical to
    that shown in Fig.~\ref{image-p}, and shows one period of the
    oscillation after the system has settled down.  The vertical axis
    shows stacked the total number of proteins that are of four
    different compound stages and in the right section of the cell
    (bottom four colors), in the middle of the cell (middle four
    colors) and in the left section of the cell (top four colors). The
    different forms of protein changes while the total number of MinD
    protein in the cell remains constant, shown by the fact that the
    very top line is horizontal.}
  \label{box-p}
\end{figure}

While Fig.~\ref{image-p} is helpful for visualizing
the process occuring in the Min system, it is does not show the
relative number of proteins in each state.  To visualize this
information, we plot in Fig.~\ref{box-p} the fraction of MinD found in
each stage of the cycle in each of three sections of the cell---top,
middle and bottom---as a function of time over one period.  The time
axis is the same as in Fig.~\ref{image-p}.  Looking at the evolution
of the latter, we see an initial high concentration of cytoplasmic
MinD:ATP, as the this protein diffuses into the polar region, followed
by a dip in this and large spike in membrane bound MinD:ATP, as the
protein clusters onto the polar cell wall, followed by a spike in the
memebrane bound MinE:MinD:ATP stage of protein, together with a slight
rise in cytoplasmic MinD:ADP as the MinD dettaches from the wall.

@@

We can see a
clear difference in period between the $5\mu$, $4\mu$, and $2\mu$
length cells.  The oscillation periods for our $4\mu$ show excellent
agreement with the simulations of Huang et all
\cite{huang2003dynamic}.  This is to be expected since we begin our
simulations with their reported wild type concentrations of MinD and
MinE proteins. The periods for the different pill lengths are shown in
the table below.  We can see a direct correltation between size and
period, as the proteins must diffuse further in a longer cell before
they accumulate on the opposite wall.
